Comparison of Fast Imaging Architectures for Large, Multi-scale, Hierarchical Aperture Arrays

### 2024-06-11 ms.tex
Update on Overleaf.

The paper is organised as follows. In \S\ref{sec:notation}, the mathematical notation describing the measurements and imaging process is established. \S\ref{sec:motivation} describes the scientific and technical motivations for this paper. \S\ref{sec:computational-cost} describes the metric of computational cost density in discovery phase space based on which we compare different imaging architectures described in \S\ref{sec:img-archs}. The multi-scale imaging architecture options for intra- and inter-station data processing are described in sections \ref{sec:inter-station-arch} and \ref{sec:inter-station-arch}, respectively. \S\ref{sec:results} presents a summary of the results, and conclusions are presented in \S\ref{sec:conclusion}.

@@

\section{Motivations}\label{sec:motivation}

This study is motivated by (a) the field of radio transient phenomena that span large ranges in time, and (b) large, modern radio interferometer arrays that access spatial scales spanning several orders of magnitude.

\subsection{Scientific: Time domain astronomy}\label{sec:time-domain-astronomy}

\subsection{Technical: Multi-scale, Hierarchical Aperture Arrays} \label{sec:multi-scale-arrays}

Modern radio arrays tend to be aperture arrays that operate interferometrically and hierarchically on multiple spatial scales. Examples of these multi-scale hierarchical aperture arrays include the Murchison Widefield Array \cite[MWA;][]{Tingay+2013}, the Low Frequency Array \cite[LOFAR;][]{vanHaarlem+2013}, the swarm of Long Wavelength Arrays \cite[LWA Swarm;][]{Dowell+2018}, and the Square Kilometre Array low \cite[SKA-low;][]{Dewdney+2009}. The specific spatial scales are the dimensions of the antenna, station (a grouping of antennas), and the array (a grouping of stations), denoted by $D_\textrm{a}$, $D_\textrm{s}$, and $D_\textrm{A}$, respectively. The number of stations in the array and the number of antennas in a station (indexed by $m$) are denoted by $N_\textrm{s}$ and $N_{\textrm{a/s}_m}$, respectively. 
% Although $N_\textrm{a/s_m}$ could vary with each station, we assume it remains the same for each station for simplicity.

@@

We have considered aperture arrays consisting of three spatial scales, but the analysis can be easily extended to other arrays consisting of even larger levels of spatial hierarchy.

@@

We consider two broad classes of two-stage fast imaging architectures. In either case, the first stage is intra-station interferometry of the element electric fields using the options of voltage beamforming (BF), E-field Parallel Imaging Correlator (EPIC), beamforming of correlations (XBF), or correlation followed by FFT (XFFT). The second stage inter-station combination of the data products from the first stage. The main difference between the two classes is in whether the second stage uses the first stage products coherently or incoherently. A schematic of these hybrid architectures and their computational cost components are illustrated in Figure~\ref{fig:hybrid-architectures-cost-schematic}. Below, we provide details on the computational cost budget for these hybrid architectures along with the merits and demerits of either classes.

>>>>

Modern radio arrays tend to be aperture arrays that operate interferometrically and hierarchically on multiple spatial scales. Examples of these multi-scale hierarchical aperture arrays include the Murchison Widefield Array \cite[MWA;][]{Tingay+2013}, the Low Frequency Array \cite[LOFAR;][]{vanHaarlem+2013}, the swarm of Long Wavelength Arrays \cite[LWA Swarm;][]{Dowell+2018}, and the Square Kilometre Array low \cite[SKA-low;][]{Dewdney+2009}.

The paper is organised as follows. In \S\ref{sec:notation}, the mathematical notation describing the measurements and imaging process is established. \S\ref{sec:multi-scale-arrays} describes the parametrisation of multi-scale aperture arrays studied in this paper. \S\ref{sec:computational-cost} describes the metric of computational cost density in discovery phase space based on which we compare different imaging architectures described in \S\ref{sec:img-archs}. The multi-scale imaging architecture options for intra- and inter-station data processing are described in sections \ref{sec:inter-station-arch} and \ref{sec:inter-station-arch}, respectively. \S\ref{sec:results} presents a summary of the results, and conclusions are presented in \S\ref{sec:conclusion}.

@@

% \section{Motivations}\label{sec:motivation}

% This study is motivated by (a) the field of radio transient phenomena that span large ranges in time, and (b) large, modern radio interferometer arrays that access spatial scales spanning several orders of magnitude.

% \subsection{Scientific: Time domain astronomy}\label{sec:time-domain-astronomy}

\section{Multi-scale, Hierarchical Aperture Arrays} \label{sec:multi-scale-arrays}

The hierarchical, multi-scale aperture array scenario we consider involves spatial scales corresponding to the dimensions of the antenna, station (a grouping of antennas), and the array (a grouping of stations), denoted by $D_\textrm{a}$, $D_\textrm{s}$, and $D_\textrm{A}$, respectively. The number of stations in the array and the number of antennas in a station (indexed by $m$) are denoted by $N_\textrm{s}$ and $N_{\textrm{a/s}_m}$, respectively. 
% Although $N_\textrm{a/s_m}$ could vary with each station, we assume it remains the same for each station for simplicity.

@@

Although we have considered aperture arrays consisting of three spatial scales, the analysis can be extended without loss of generality to arrays consisting of even larger levels of spatial hierarchy.

@@

We consider two broad classes of two-stage fast imaging architectures. In both classes, the first stage is intra-station interferometry of the element electric fields using the options of voltage beamforming (BF), E-field Parallel Imaging Correlator (EPIC), beamforming of correlations (XBF), or correlation followed by FFT (XFFT). The second stage inter-station combination of the data products from the first stage. The main difference between the two classes is in whether the second stage uses the first stage products coherently or incoherently. A schematic of these hybrid architectures and their computational cost components are illustrated in Figure~\ref{fig:hybrid-architectures-cost-schematic}. Below, we provide details on the computational cost budget for these hybrid architectures along with the merits and demerits of either classes.
